The tools of valuation empower the investor to judge stocks in a manner that is detached from behavioral biases and therefore from market sentiment. Valuation emphasizes companies' fundamentals and financial health.

By relying on valuation -- regardless of an active or passive investing style -- we reject the Efficient Market Hypothesis (EMH) in at least its strong and semi-strong forms.

\subsection{DCF (Deterministic)}

\subsubsection{Discounted Cash Flow Foundation}

Let us derive a stock's value per share based on discounted free cash flow and regard this value as intrinsic.

Interest, e.g., on investment streams, compounds to form the future value:
\[
\text{FutureValue}_n = \sum_{t=0}^{n} \text{InvestedCash}_t \cdot (1 + \text{InterestRate})^{n - t}
\]

To calculate the present value, e.g. of a business, we undo the compounding effect. The intrinsic value of a financial asset is the discounted value of its future cash flows:
\[
\text{PresentValue} = \sum_{t=1}^{n} \frac{\text{CashFlow}_t}{(1 + \text{DiscountRate})^t}
\]

One can go the free-cash-flow-to-equity (FCFE) and/or the free-cash-flow-to-firm (FCFF) route. It will depend on the company's circumstances to which to give modeling precedence. FCFE should lead to a more insightful intrinsic value per share (IVPS), if the debt policy is known and reliable, and FCFF should do so, if the debt structure is uncertain or complex.

\subsubsection{Cash Flow Mechanics}

\paragraph{Free Cash Flow to Equity}

$$
\text{FCFE} = \text{NI} - (1 - \text{TDR}) \cdot (\text{CAPX} - \text{D} - \Delta \text{WC})
$$

\begin{itemize}
  \item[] $\text{FCFE}$: free cash flow to equity
  \item[] $\text{NI}$: net income
  \item[] $\text{TDR}$: target debt ratio
  \item[] $\text{CAPX}$: capital expenditure
  \item[] $\text{D}$: depreciation \& amortization
  \item[] $\Delta \text{WC} = \Delta \text{CA} - \Delta \text{CL}$: change in working capital
  \item[] $\text{CA}$: current assets
  \item[] $\text{CL}$: current liabilities
\end{itemize}

\paragraph{Free Cash Flow to Firm}

$$
\text{FCFF} = \text{EBIT} \cdot (1 - \text{CTR}) + \text{D} - \text{CAPX} - \Delta \text{WC}
$$

\begin{itemize}
  \item[] $\text{FCFF}$: free cash flow to firm
  \item[] $\text{EBIT}$: earnings before interest and taxes
  \item[] $\text{CTR}$: corporate tax rate
\end{itemize}

\subsubsection{Cost of Capital}

\paragraph{Levered Beta}

$$
\beta_{\text{L}} = \beta_{\text{U}} \cdot \left(1 + (1 - \text{CTR}) \cdot \frac{\text{MVB}}{\text{MVE}}\right)
$$

\begin{itemize}
  \item[] $\beta_{\text{L}}$: levered beta
  \item[] $\beta_{\text{U}}$: unlevered beta
  \item[] $\text{MVB}$: market value of debt
  \item[] $\text{MVE}$: market value of equity
\end{itemize}

\paragraph{Cost of Equity (CAPM)}

$$
\text{CE} = \text{RFR} + \beta_{\text{L}} \cdot \text{ERP}
$$

\begin{itemize}
  \item[] $\text{CE}$: cost of equity
  \item[] $\text{RFR}$: risk-free rate
  \item[] $\text{ERP}$: equity risk premium
\end{itemize}

\paragraph{Currency-Adjusted Equity Risk Premium}

$$
\text{ERPT} = (1 + \text{ERPS}) \cdot \left( \frac{1 + \text{IT}}{1 + \text{IS}} \right) - 1
$$

\begin{itemize}
  \item[] $\text{ERPT}$: equity risk premium of target currency
  \item[] $\text{ERPS}$: equity risk premium of source currency
  \item[] $\text{IT}$: inflation of target
  \item[] $\text{IS}$: inflation of source
\end{itemize}

\paragraph{Weighted Average Cost of Capital}

$$
\text{WACC} = \frac{\text{MVE}}{\text{MVE} + \text{MVB}} \cdot \text{CE} + \frac{\text{MVB}}{\text{MVE} + \text{MVB}} \cdot \text{CB} \cdot (1 - \text{CTR})
$$

\begin{itemize}
  \item[] $\text{WACC}$: weighted average cost of capital
  \item[] $\text{CB}$: pre-tax cost of debt
\end{itemize}

\subsubsection{Growth Rate Estimation}

\paragraph{FCFE Growth Rate}

$$
\begin{aligned}
\text{FCFEGR} &= \text{ROE} \cdot \text{FCFERR} \\
&= \left( \frac{\text{NI}}{\text{BVE}} \right) \cdot \left( 1 - \frac{\text{DP}}{\text{NI}} \right)
\end{aligned}
$$

\begin{itemize}
  \item[] $\text{FCFEGR}$: FCFE growth rate estimate
  \item[] $\text{ROE}$: return on equity
  \item[] $\text{FCFERR}$: FCFE retention ratio
  \item[] $\text{BVE}$: book value of equity
  \item[] $\text{DP}$: dividends paid
\end{itemize}

\paragraph{FCFF Growth Rate}

$$
\begin{aligned}
\text{FCFFGR} &= \text{ROIC} \cdot \text{FCFFRR} \\
&= \left( \frac{\text{NOPAT}}{\text{IC}} \right) \cdot \left( \frac{\text{CAPX} - \text{D} + \Delta \text{WC}}{\text{NOPAT}} \right) \\
&= \left( \frac{\text{EBIT} \cdot (1 - \text{ETR})}{\text{IC}} \right) \cdot \left( \frac{\text{CAPX} - \text{D} + \Delta \text{WC}}{\text{EBIT} \cdot (1 - \text{ETR})} \right)
\end{aligned}
$$

\begin{itemize}
  \item[] $\text{FCFFGR}$: FCFF growth rate estimate
  \item[] $\text{ROIC}$: return on invested capital
  \item[] $\text{FCFFRR}$: FCFF reinvestment rate
  \item[] $\text{NOPAT}$: net operating profit after taxes
  \item[] $\text{IC}$: invested capital
  \item[] $\text{ETR}$: effective tax rate
\end{itemize}

\subsubsection{Present Value Calculation}

\paragraph{FCFE-Based Valuation}

$$
\begin{aligned}
\text{PVE} = \text{PVFCFE} &+ \text{PVTVFCFE} \\
= \left( \sum_{t=1}^{h} \frac{\text{FCFE}_t}{(1 + \text{CE})^t} \right) &+ \text{PVTVFCFE} \\
= \left( \sum_{t=1}^{h} \frac{\text{FCFE}_0 \cdot (1+\text{FCFEGR})^t}{(1+\text{CE})^t} \right) &+ \text{PVTVFCFE} \\
= \text{PVFCFE} &+ \left( \sum_{t=h+1}^{\infty} \frac{\text{FCFE}_t}{(1 + \text{CE})^t} \right) \\
= \text{PVFCFE} &+ \left( \sum_{t=h+1}^{\infty} \frac{\text{FCFE}_h \cdot (1+\text{TGR})^{t-h}}{(1+\text{CE})^t} \right) \\
= \text{PVFCFE} &+ \left( \sum_{t=h+1}^{\infty} \frac{\text{FCFE}_h}{(1+\text{CE})^h} \cdot \left( \frac{1+\text{TGR}}{1+\text{CE}} \right)^{t-h} \right) \\
= \text{PVFCFE} &+ \left( \frac{\text{FCFE}_h}{(1+\text{CE})^h} \sum_{k=1}^{\infty} \left( \frac{1+\text{TGR}}{1+\text{CE}} \right)^k \right) \\
\overset{\text{GGM}}{=} \text{PVFCFE} &+ \left( \frac{\text{FCFE}_h}{(1+\text{CE})^h} \cdot \frac{\frac{1+\text{TGR}}{1+\text{CE}}}{1 - \frac{1+\text{TGR}}{1+\text{CE}}} \right) \\
= \text{PVFCFE} &+ \left( \frac{\text{FCFE}_h}{(1+\text{CE})^h} \cdot \frac{1+\text{TGR}}{\text{CE} - \text{TGR}} \right) \\
= \text{PVFCFE} &+ \left( \frac{\text{FCFE}_h \cdot (1+\text{TGR})}{(\text{CE} - \text{TGR})} \cdot \frac{1}{(1+\text{CE})^h} \right) \\
= \text{PVFCFE} &+ \left(\frac{\text{FCFE}_{h+1}}{(\text{CE} - \text{TGR})} \cdot \frac{1}{(1 + \text{CE})^h} \right) \\
= \text{PVFCFE} &+ \left(\frac{\text{TVFCFE}}{(1 + \text{CE})^h} \right) \\
= \left( \sum_{t=1}^{h} \frac{\text{FCFE}_0 \cdot (1+\text{FCFEGR})^t}{(1+\text{CE})^t} \right) &+ \left(\frac{\text{FCFE}_h \cdot (1+\text{TGR})}{(\text{CE} - \text{TGR}) \cdot (1 + \text{CE})^h} \right)
\end{aligned}
$$

\begin{itemize}
  \item[] $\text{PVE}$: present value of equity
  \item[] $\text{PVFCFE}$: present value of free cash flow to equity
  \item[] $\text{PVTVFCFE}$: present value of terminal value of FCFE
  \item[] $h$: growth forecast horizon (before terminal growth rate into perpetuity)
  \item[] $\text{GGM}$: Gordon growth model (closed-form solution of infinite geometric series)
  \item[] $\text{TGR}$: terminal growth rate
  \item[] $\text{TVFCFE}$: terminal value of FCFE
\end{itemize}

\paragraph{FCFF-Based Valuation}

$$
\begin{aligned}
\text{PVF} = \text{PVFCFF} &+ \text{PVTVFCFF} \\
= \left( \sum_{t=1}^{h} \frac{\text{FCFF}_t}{(1 + \text{WACC})^t} \right) &+ \text{PVTVFCFF} \\
= \left( \sum_{t=1}^{h} \frac{\text{FCFF}_0 \cdot (1+\text{FCFFGR})^t}{(1+\text{WACC})^t} \right) &+ \text{PVTVFCFF} \\
= \text{PVFCFF} &+ \left( \sum_{t=h+1}^{\infty} \frac{\text{FCFF}_t}{(1 + \text{WACC})^t} \right) \\
= \text{PVFCFF} &+ \left( \sum_{t=h+1}^{\infty} \frac{\text{FCFF}_h \cdot (1+\text{TGR})^{t-h}}{(1+\text{WACC})^t} \right) \\
= \text{PVFCFF} &+ \left( \sum_{t=h+1}^{\infty} \frac{\text{FCFF}_h}{(1+\text{WACC})^h} \cdot \left( \frac{1+\text{TGR}}{1+\text{WACC}} \right)^{t-h} \right) \\
= \text{PVFCFF} &+ \left( \frac{\text{FCFF}_h}{(1+\text{WACC})^h} \sum_{k=1}^{\infty} \left( \frac{1+\text{TGR}}{1+\text{WACC}} \right)^k \right) \\
\overset{\text{GGM}}{=} \text{PVFCFF} &+ \left( \frac{\text{FCFF}_h}{(1+\text{WACC})^h} \cdot \frac{\frac{1+\text{TGR}}{1+\text{WACC}}}{1 - \frac{1+\text{TGR}}{1+\text{WACC}}} \right) \\
= \text{PVFCFF} &+ \left( \frac{\text{FCFF}_h}{(1+\text{WACC})^h} \cdot \frac{1+\text{TGR}}{\text{WACC} - \text{TGR}} \right) \\
= \text{PVFCFF} &+ \left( \frac{\text{FCFF}_h \cdot (1+\text{TGR})}{(\text{WACC} - \text{TGR})} \cdot \frac{1}{(1+\text{WACC})^h} \right) \\
= \text{PVFCFF} &+ \left( \frac{\text{FCFF}_{h+1}}{(\text{WACC} - \text{TGR})} \cdot \frac{1}{(1 + \text{WACC})^h} \right) \\
= \text{PVFCFF} &+ \left( \frac{\text{TVFCFF}}{(1 + \text{WACC})^h} \right) \\
= \left( \sum_{t=1}^{h} \frac{\text{FCFF}_0 \cdot (1+\text{FCFFGR})^t}{(1+\text{WACC})^t} \right) &+ \left( \frac{\text{FCFF}_h \cdot (1+\text{TGR})}{(\text{WACC} - \text{TGR}) \cdot (1 + \text{WACC})^h} \right)
\end{aligned}
$$

\begin{itemize}
  \item[] $\text{PVF}$: present value of the firm
  \item[] $\text{PVFCFF}$: present value of free cash flow to firm
  \item[] $\text{PVTVFCFF}$: present value of terminal value of FCFF
  \item[] $\text{TVFCFF}$: terminal value of FCFF
\end{itemize}

\paragraph{Reconciliation}

$$
\text{PVE} \approx \text{PVF} - \text{MVB} + \text{CNOA}
$$

\begin{itemize}
  \item[] $\text{CNOA}$: cash and non-operating assets
\end{itemize}

\paragraph{Intrinsic Value Per Share}

$$
\text{IVPS} = \frac{\text{PVE}}{\text{SO}}
$$

\begin{itemize}
  \item[] $\text{IVPS}$: intrinsic value per share
  \item[] $\text{SO}$: shares outstanding
\end{itemize}

\subsubsection{Interpretation}

Having derived an $\text{IVPS}$, it may serve as a reference point when judging the market price. Supposing that, ideally, both (rather than just one of) the FCFE and FCFF based IVPS have been modeled well and are therefore reliable indicators, we may interpret the nine resulting constellations as follows.

\renewcommand{\arraystretch}{1.8}

\noindent
\begin{tabular}{>{\raggedright\arraybackslash}p{0.15\textwidth}
                >{\raggedright\arraybackslash}p{0.28\textwidth}
                >{\raggedright\arraybackslash}p{0.28\textwidth}
                >{\raggedright\arraybackslash}p{0.28\textwidth}}

\textbf{}
& \textbf{FCFE $>$ Price}
& \textbf{FCFE $\approx$ Price}
& \textbf{FCFE $<$ Price} \\

\textbf{FCFF $>$ Price}
& \signalstrongbuy{} The firm's assets and operations generate more value than what is priced in by the market, and equity holders retain it — low leverage or efficient debt structure.
& \signalbuy{} The firm's assets and operations are underpriced, but excess value is absorbed by debt or reinvestment, leaving equity fairly valued.
& \signalcaution{} The firm's assets and operations are underpriced, but debt or reinvestment absorbs most cash flows — equity claims more than it economically receives. \\

\textbf{FCFF $\approx$ Price}
& \signalbuy{} The firm's assets and operations are fairly priced, but equity captures a disproportionately large share — market underprices the equity upside.
& \signalhold{} The present value of free cash flows — to the firm (before payments to debt holders) and to equity (after them) — is consistent with market prices; no mispricing is evident.
& \signalspeculative{} The business generates enough pre-financing cash flow to justify its market price, but equity holders retain too little after payments to debt holders. \\

\textbf{FCFF $<$ Price}
& \signalcaution{} The business is overvalued, but equity appears cheap due to temporarily favorable debt terms — value may be unstable under a leveraged structure.
& \signalspeculative{} Equity is fairly priced, but depends on cash flows from a business generating less than what its market price would suggest — any decline in operations could undermine equity value.
& \signalavoid{} There isn't sufficient cash flow to the business or the equity for the fundamentals to justify the high market price. \\
\end{tabular}

\subsubsection{Sector-Specific Models}

The standard FCFE/FCFF framework breaks down for three categories of companies whose economics fundamentally differ from the general corporate model. For each, we derive a purpose-built valuation and sensitivity analysis.

\paragraph{Bank Valuation (Excess Return Model)}

For banks, interest expense is an operating cost (not a financing cost), and deposits -- the primary source of debt -- are the raw material of the business. FCFE and FCFF are therefore not meaningful. Instead, we value banks through the excess return model:

$$
\text{FV}_{\text{bank}} = \text{BVE} + \sum_{t=1}^{h} \frac{(\text{ROE}_t - \text{CE}) \cdot \text{BVE}_t}{(1 + \text{CE})^t} + \frac{(\text{ROE}_{\infty} - \text{CE}) \cdot \text{BVE}_{h}}{(\text{CE} - \text{TGR}) \cdot (1 + \text{CE})^h}
$$

where ROE mean-reverts toward the cost of equity at rate $\lambda$:

$$
\text{ROE}_{t+1} = \text{ROE}_t + \lambda \cdot (\text{CE} - \text{ROE}_t)
$$

\begin{itemize}
  \item[] $\text{BVE}$: book value of equity
  \item[] $\text{ROE}_t$: return on equity in year $t$
  \item[] $\text{ROE}_{\infty}$: terminal ROE (cost of equity plus sustainable spread)
  \item[] $\lambda$: mean-reversion speed (typically $\approx 0.15$)
\end{itemize}

The model's implied P/BV follows a Gordon-growth-style identity:

$$
\frac{P}{B} = \frac{\text{ROE} - g}{\text{CE} - g}
$$

When $\text{ROE} = \text{CE}$, the stock should trade at book value. The sensitivity analysis sweeps ROE ($5\text{--}25\%$), cost of equity ($6\text{--}12\%$), and sustainable growth ($0\text{--}6\%$).

\paragraph{Insurance Valuation (Float-Based Model)}

Insurers collect premiums before paying claims, creating a pool of investable capital called \emph{float}. If underwriting is profitable (combined ratio $< 100\%$), this float is obtained at negative cost -- policyholders effectively pay the insurer to hold their money.

$$
\text{FV}_{\text{ins}} = \text{BVE} + \text{UW}_{\text{PV}} + \text{Float}_{\text{PV}}
$$

where the present value of underwriting profits is:

$$
\text{UW}_{\text{PV}} = \sum_{t=1}^{h} \frac{P_t \cdot (1 - \text{CR}_t)}{(1 + \text{CE})^t} + \frac{P_{h} \cdot (1 - \text{CR}_{\infty})}{(\text{CE} - \text{TGR}) \cdot (1 + \text{CE})^h}
$$

and the float value is a perpetuity of net investment income:

$$
\text{Float}_{\text{PV}} = \frac{F \cdot (\text{IY} - \text{CoF})}{\text{CE} - \text{TGR}}
$$

\begin{itemize}
  \item[] $P_t$: premiums earned in year $t$
  \item[] $\text{CR}_t$: combined ratio (loss ratio + expense ratio)
  \item[] $F$: float (unearned premiums + loss reserves)
  \item[] $\text{IY}$: investment yield on float
  \item[] $\text{CoF}$: cost of float ($= \text{CR} - 1$ if $\text{CR} > 1$; 0 otherwise)
\end{itemize}

The combined ratio mean-reverts toward the industry average ($\approx 98\%$). The sensitivity analysis sweeps combined ratio ($85\text{--}110\%$), investment yield ($1\text{--}6\%$), and cost of equity ($6\text{--}12\%$).

\paragraph{Oil \& Gas E\&P Valuation (NAV Model)}

Exploration and production companies are valued on the net asset value of their proven reserves, projected using a decline-curve model:

$$
\text{FV}_{\text{og}} = \frac{1}{2}\left(\text{NAV} + \text{PV}_{10}\right)
$$

where

$$
\text{NAV} = \sum_{t=1}^{T_{\text{res}}} \frac{Q_t \cdot (\text{NB}_t) \cdot (1 - \tau)}{(1 + r)^t} - D
$$

\begin{itemize}
  \item[] $Q_t = Q_0 \cdot (1 - d)^t$: production in year $t$ with decline rate $d$
  \item[] $\text{NB} = p_{\text{oil}} \cdot \alpha + p_{\text{gas}} \cdot 6 \cdot (1 - \alpha) - c_{\text{lift}}$: netback per BOE
  \item[] $\alpha$: oil production fraction
  \item[] $c_{\text{lift}}$: lifting cost per BOE
  \item[] $T_{\text{res}}$: reserve life (capped at 30 years)
  \item[] $\text{PV}_{10}$: SEC-standard valuation at $r = 10\%$
  \item[] $D$: net debt
\end{itemize}

The blended fair value averages NAV (company-specific discount rate) and PV-10 (standard 10\% rate). The sensitivity analysis sweeps oil price (\$40--\$120/bbl), lifting cost (\$5--\$25/BOE), and discount rate ($6\text{--}14\%$).

\subsubsection{Sensitivity Analysis}

Each valuation model produces a four-panel sensitivity plot: three line charts showing fair value as a function of one swept parameter (with the market price as a reference line and the implied parameter value marked), and a tornado diagram ranking parameters by their impact on fair value.

\renewcommand{\arraystretch}{1.4}

\noindent
\begin{tabular}{lll}
\textbf{Model} & \textbf{Parameter Axes} & \textbf{Output} \\
\hline
Standard DCF & Growth rate, Discount rate, Terminal growth & FCFE \& FCFF IVPS \\
Bank & ROE, Cost of equity, Sustainable growth & Fair value \\
Insurance & Combined ratio, Investment yield, Cost of equity & Fair value \\
Oil \& Gas & Oil price, Lifting cost, Discount rate & Blended NAV \\
\end{tabular}

\vspace{0.5em}

The implied parameter value at which fair value equals the market price reveals what the market is pricing in. For banks, this is the implied ROE; for insurers, the implied combined ratio; for O\&G companies, the implied oil price. Deviation between the actual and implied parameter indicates the degree to which the market under- or over-prices the asset.

\newpage\subsection{DCF (Probabilistic)}

\subsubsection{Motivation}

The deterministic DCF model relies on point estimates for key inputs (growth rates, discount rates, tax rates). However, these parameters are inherently uncertain. Probabilistic DCF addresses this by:

\begin{enumerate}
  \item Treating financial statement line items as random variables
  \item Propagating uncertainty through the cash flow projection
  \item Generating a distribution of intrinsic values rather than a single point estimate
  \item Quantifying valuation confidence via probability metrics
\end{enumerate}

\subsubsection{Input Sampling}

\paragraph{Historical Distributions}

The following variables are sampled independently in each simulation iteration $i = 1, \ldots, N$ (typically $N = 5000$):

\begin{itemize}
  \item Net income: $\text{NI}^{(i)}$
  \item Earnings before interest and taxes: $\text{EBIT}^{(i)}$
  \item Capital expenditures: $\text{CAPX}^{(i)}$
  \item Depreciation \& amortization: $\text{D}^{(i)}$
  \item Current assets: $\text{CA}^{(i)}$
  \item Current liabilities: $\text{CL}^{(i)}$
\end{itemize}

For each variable $X$, we compute empirical mean $\mu_X$ and standard deviation $\sigma_X$ from the historical time series (typically 4--8 quarters).

\paragraph{Sampling Distributions}

We choose the sampling distribution based on the sign and scale of the variable:

\begin{itemize}
  \item If $\mu_X > 0$, the variable is sampled from a log-normal distribution to ensure positivity:
  \[
  \log\text{-mean} = \log(\max(\mu_X, 10^{-3})), \quad \log\text{-std} = \frac{\sigma_X}{\mu_X}
  \]
  \[
  X^{(i)} \sim \text{LogNormal}(\log\text{-mean}, \log\text{-std})
  \]

  \item If $\mu_X \leq 0$, the variable falls back to a normal distribution:
  \[
  X^{(i)} \sim \mathcal{N}(\mu_X, \sigma_X^2)
  \]
\end{itemize}

No additional prior smoothing is applied to these inputs. Their randomness enters directly into the cash flow projection models.

\paragraph{Outlier Squashing}

To reduce the impact of extreme outliers, selected variables like $\text{NI}$ and $\text{EBIT}$ are post-processed using a squashing function:
\[
x \mapsto
\begin{cases}
x, & x \leq \tau \\
\tau + \log(1 + (x - \tau)), & x > \tau
\end{cases}
\]
where the threshold $\tau$ is:
\[
\tau = \min\{3 \cdot \max(\text{historical values}), \, 0.3 \cdot \text{MarketCap}\}
\]

This logarithmic damping prevents unrealistic extrapolation while preserving the shape of the distribution.

\subsubsection{Growth Rate Sampling}

\paragraph{Fundamental Growth Identity}

Two growth rates are sampled in each simulation:
\begin{itemize}
  \item FCFE growth rate: $g_{\text{FCFE}}$
  \item FCFF growth rate: $g_{\text{FCFF}}$
\end{itemize}

Both follow the fundamental identity:
\[
g = \text{return} \times \text{reinvestment rate}
\]

For FCFE:
\begin{align*}
g_{\text{FCFE}} &= \text{ROE} \times (1 - \text{payout ratio}) \\
&= \frac{\text{NI}}{\text{BVE}} \times \left(1 - \frac{\text{DP}}{\text{NI}}\right)
\end{align*}

For FCFF:
\begin{align*}
g_{\text{FCFF}} &= \text{ROIC} \times \text{reinvestment rate} \\
&= \frac{\text{NOPAT}}{\text{IC}} \times \frac{\text{CAPX} - \text{D} + \Delta \text{WC}}{\text{NOPAT}}
\end{align*}

where $\text{NOPAT} = 0.75 \cdot \text{EBIT}$ (approximating $(1 - \text{CTR}) \cdot \text{EBIT}$ with a fixed $25\%$ tax rate).

\paragraph{Bayesian Smoothing with Sector Priors}

Each return and reinvestment term is blended as:
\[
\text{final} = 0.5 \cdot \text{empirical} + 0.5 \cdot \text{prior}
\]

Sector-specific priors are sampled from:
\begin{itemize}
  \item ROE and ROIC: $\text{Beta}(2, 5)$ scaled to $[0, 0.4]$
  \item Retention and reinvestment rates: $\text{Beta}(2, 2)$ scaled to $[0, 1]$
\end{itemize}

This regularization prevents overfitting to recent historical data while incorporating industry-level information.

\paragraph{Growth Rate Distribution}

The blended product gives a per-year estimate. Growth rates are then sampled from a normal distribution using the historical mean and standard deviation:
\[
g^{(i)} \sim \mathcal{N}(\mu_g, \sigma_g^2)
\]

and clipped to user-defined bounds:
\[
g^{(i)} \in [g_{\min}, g_{\max}]
\]

Typical bounds are $g_{\min} = -0.05$ (allowing slight contraction) and $g_{\max} = 0.15$ (capping growth at 15\%).

This ensures simulated growth is consistent with both financial fundamentals and prior regularization.

\subsubsection{Stochastic Discount Rates}

In addition to stochastic growth rates, the discount rate components (risk-free rate, beta, and equity risk premium) may also be sampled to capture uncertainty in the cost of capital.

\paragraph{Discount Rate Component Sampling}

For each simulation $i$, the following components are sampled from normal distributions:

\begin{itemize}
  \item Risk-free rate:
  \[
  \text{RFR}^{(i)} \sim \mathcal{N}(\text{RFR}_{\text{base}}, \sigma_{\text{RFR}}^2)
  \]
  with typical volatility $\sigma_{\text{RFR}} = 0.005$ (50 basis points).

  \item Leveraged beta:
  \[
  \beta_{\text{L}}^{(i)} \sim \mathcal{N}(\beta_{\text{L,base}}, \sigma_{\beta}^2)
  \]
  with typical volatility $\sigma_{\beta} = 0.10$.

  \item Equity risk premium:
  \[
  \text{ERP}^{(i)} \sim \mathcal{N}(\text{ERP}_{\text{base}}, \sigma_{\text{ERP}}^2)
  \]
  with typical volatility $\sigma_{\text{ERP}} = 0.01$ (100 basis points).
\end{itemize}

\paragraph{Cost of Capital Recomputation}

Using the sampled components, the cost of equity for simulation $i$ is:
\[
\text{CE}^{(i)} = \text{RFR}^{(i)} + \beta_{\text{L}}^{(i)} \cdot \text{ERP}^{(i)}
\]

Similarly, the weighted average cost of capital is recomputed:
\[
\text{WACC}^{(i)} = \frac{\text{MVE}}{\text{MVE} + \text{MVB}} \cdot \text{CE}^{(i)} + \frac{\text{MVB}}{\text{MVE} + \text{MVB}} \cdot \text{CB} \cdot (1 - \text{CTR})
\]

This stochastic treatment of discount rates significantly increases the dispersion of the intrinsic value distribution, providing a more realistic quantification of valuation uncertainty. The feature is controlled by the \texttt{use\_stochastic\_discount\_rates} configuration flag.

\subsubsection{Cash Flow Projection}

\paragraph{Explicit Forecast Period}

For each simulation $i$, cash flows are projected over $h$ years (typically $h = 5$) using the sampled inputs and growth rate.

\paragraph{Time-Varying Growth Rates}

Rather than applying a constant growth rate throughout the projection period, the model implements exponential mean reversion toward the terminal growth rate. This prevents unrealistic perpetual high-growth assumptions.

For each simulation $i$ and forecast year $t \in \{1, \ldots, h\}$, the growth rate decays according to:
\[
g_t^{(i)} = g_{\text{terminal}} + \left(g_0^{(i)} - g_{\text{terminal}}\right) \cdot e^{-\lambda t}
\]

where:
\begin{itemize}
  \item $g_0^{(i)}$ is the initial sampled growth rate (FCFE or FCFF growth from the fundamental identity)
  \item $g_{\text{terminal}}$ is the long-run terminal growth rate (typically 2--3\%)
  \item $\lambda \in [0, 1]$ is the mean reversion speed parameter (typical value: 0.3)
  \item $t$ is the year index within the projection period
\end{itemize}

\textbf{Interpretation of $\lambda$:}
\begin{itemize}
  \item $\lambda = 0$: No reversion, constant growth $g_t^{(i)} = g_0^{(i)}$ for all $t$
  \item $\lambda = 0.3$: Moderate reversion (recommended), growth decays smoothly over 5--7 years
  \item $\lambda = 1.0$: Fast reversion, growth approaches terminal rate within 3--4 years
\end{itemize}

This time-varying formulation yields more conservative valuations than constant-growth models, as high initial growth rates gradually moderate to sustainable long-term levels. The feature is controlled by the \texttt{use\_time\_varying\_growth} configuration flag.

\paragraph{Cash Flow Computation}

Using the time-varying growth rates $g_t^{(i)}$, the projected cash flows are:

For FCFE:
\[
\text{FCFE}_t^{(i)} = \text{NI}_t^{(i)} + \text{D}_t^{(i)} - \text{CAPX}_t^{(i)} - \Delta \text{WC}_t^{(i)} + \text{NetBorrowing}_t^{(i)}
\]
where $\text{NetBorrowing}_t^{(i)} = \text{TDR} \cdot \text{Reinvestment}_t^{(i)}$

For FCFF:
\[
\text{FCFF}_t^{(i)} = \text{NOPAT}_t^{(i)} + \text{D}_t^{(i)} - \text{CAPX}_t^{(i)} - \Delta \text{WC}_t^{(i)}
\]
with $\text{NOPAT}_t^{(i)} = (1 - \text{CTR}) \cdot \text{EBIT}_t^{(i)}$

\paragraph{Terminal Value}

The terminal value uses a Gordon growth model:
\[
\text{TV}^{(i)} = \frac{\text{CashFlow}_h^{(i)} \cdot (1 + g_{\text{terminal}})}{r - g_{\text{terminal}}}
\]

where:
\begin{itemize}
  \item $r = \text{CE}$ for FCFE or $r = \text{WACC}$ for FCFF
  \item $g_{\text{terminal}}$ is the terminal growth rate (typically 2--3\%)
\end{itemize}

To avoid division by zero or unstable extrapolation, the denominator is lower-bounded:
\[
r - g_{\text{terminal}} \geq 0.01
\]

The terminal value is then discounted to present:
\[
\text{PV}_{\text{terminal}}^{(i)} = \frac{\text{TV}^{(i)}}{(1 + r)^h}
\]

\paragraph{Present Value}

The present value of cash flows is:
\[
\text{PV}_{\text{years}}^{(i)} = \sum_{t=1}^h \frac{\text{CashFlow}_t^{(i)}}{(1 + r)^t}
\]

Final present value of the firm:
\[
\text{PV}^{(i)} = \text{PV}_{\text{years}}^{(i)} + \text{PV}_{\text{terminal}}^{(i)}
\]

For FCFF, market value of debt per share is subtracted to get the per-share equity value:
\[
\text{IVPS}_{\text{FCFF}}^{(i)} = \frac{\text{PV}_{\text{FCFF}}^{(i)} - \text{MVB} + \text{CNOA}}{\text{SO}}
\]

\subsubsection{Statistical Analysis}

\paragraph{Simulation Matrix}

After $N$ simulations, we obtain an intrinsic value distribution for each asset and method (FCFE, FCFF):
\[
\{V^{(1)}, V^{(2)}, \ldots, V^{(N)}\}
\]

\paragraph{Kernel Density Estimation}

To estimate the probability density of intrinsic value, we apply a Gaussian kernel density estimator (KDE):
\[
\hat{f}(v) = \frac{1}{Nh} \sum_{i=1}^N K\left(\frac{v - V^{(i)}}{h}\right)
\]
where $K(\cdot)$ is the Gaussian kernel:
\[
K(u) = \frac{1}{\sqrt{2\pi}} e^{-\frac{u^2}{2}}
\]
and $h$ is the bandwidth selected using Scott's rule:
\[
h = N^{-1/5} \cdot \hat{\sigma}
\]
where $\hat{\sigma}$ is the sample standard deviation of $\{V^{(i)}\}$.

\paragraph{Value Surplus View}

Each simulation sample is transformed into a percentage surplus over market price:
\[
\text{Surplus}^{(i)} = \frac{V^{(i)} - P}{P}
\]

We then apply KDE on the transformed values to estimate the distribution of relative mispricing. This highlights the extent to which the simulated valuations exceed or fall short of the current price.

From the KDE and simulation samples, we extract:
\begin{itemize}
  \item $\mathbb{P}[\text{Surplus} > 0]$: probability the asset is undervalued
  \item $\mathbb{P}[\text{Surplus} < 0]$: probability it is overvalued
\end{itemize}

These quantities are calculated empirically via:
\[
\frac{1}{N} \sum_{i=1}^N \mathbb{I}[V^{(i)} > P]
\quad \text{and} \quad
\frac{1}{N} \sum_{i=1}^N \mathbb{I}[V^{(i)} < P]
\]

\subsubsection{Multi-Asset Portfolio Frontiers}

To explore portfolio-level valuation outcomes, we simulate many random portfolios over the set of assets.

\paragraph{Simulation Matrix}

Let $n$ be the number of assets and $S^{(j)}_i$ the $j$-th simulation of value surplus (as a percentage of market price) for asset $i$. The simulation matrix is:
\[
\mathbf{S} \in \mathbb{R}^{N \times n}, \quad \text{with } S_{ji} = \frac{V^{(j)}_i - P_i}{P_i}
\]

\paragraph{Random Portfolio Generation}

We generate $M$ random portfolio weight vectors $\mathbf{w}^{(k)} \in \mathbb{R}^n$ (typically $M = 5000$) using a Dirichlet distribution over the simplex:
\[
\mathbf{w}^{(k)} \sim \text{Dir}(\alpha \mathbf{1}), \quad \alpha = 1
\]

This ensures:
\[
\sum_{i=1}^n w^{(k)}_i = 1, \quad w^{(k)}_i \geq 0 \quad \forall i
\]

and provides uniform sampling over the weight simplex.

\paragraph{Portfolio Return Distribution}

For each portfolio $k$ and simulation $j$, the portfolio surplus is:
\[
R^{(k)}_j = \sum_{i=1}^n w^{(k)}_i S_{ji} = (\mathbf{w}^{(k)})^\top \mathbf{S}_j
\]

where $\mathbf{S}_j$ is the $j$-th row of the simulation matrix.

\paragraph{Portfolio Statistics}

For each portfolio $k$:

\begin{itemize}
  \item \textbf{Expected return} (mean surplus):
  \[
  \mu^{(k)} = \frac{1}{N} \sum_{j=1}^N R^{(k)}_j
  \]

  \item \textbf{Standard deviation} (valuation risk):
  \[
  \sigma^{(k)} = \sqrt{\frac{1}{N-1} \sum_{j=1}^N (R^{(k)}_j - \mu^{(k)})^2}
  \]

  Alternatively, using the covariance matrix:
  \[
  \sigma^{(k)} = \sqrt{(\mathbf{w}^{(k)})^\top \mathbf{\Sigma} \mathbf{w}^{(k)}}
  \]
  where $\mathbf{\Sigma}$ is the empirical covariance of $\mathbf{S}$.

  \item \textbf{Probability of loss}:
  \[
  p_{\text{loss}}^{(k)} = \frac{1}{N} \sum_{j=1}^N \mathbb{I}[R^{(k)}_j < 0]
  \]

  \item \textbf{Downside deviation} (Sortino denominator):
  \[
  \sigma_{\text{down}}^{(k)} = \sqrt{\frac{1}{N} \sum_{j=1}^N \min(0, R^{(k)}_j)^2}
  \]

  \item \textbf{Conditional Value-at-Risk} (CVaR at 95\%):
  Let $\alpha = 0.05$. Sort the portfolio returns: $R^{(k)}_{(1)} \leq R^{(k)}_{(2)} \leq \cdots \leq R^{(k)}_{(N)}$.

  The VaR is:
  \[
  \text{VaR}_{95}^{(k)} = R^{(k)}_{(\lfloor \alpha N \rfloor)}
  \]

  The CVaR is:
  \[
  \text{CVaR}_{95}^{(k)} = \frac{1}{\lfloor \alpha N \rfloor} \sum_{j=1}^{\lfloor \alpha N \rfloor} R^{(k)}_{(j)}
  \]

  \item \textbf{Maximum drawdown}:
  Compute cumulative returns over the simulation series and measure the largest peak-to-trough decline.
\end{itemize}

\paragraph{Efficient Frontier Plots}

Six distinct frontier views are produced, each optimizing a different risk-return trade-off:

\begin{enumerate}
  \item \textbf{Mean vs Standard Deviation} ($\mu$--$\sigma$):
  Mean-variance analysis. Identifies:
  \begin{itemize}
    \item Minimum variance portfolio: $\mathbf{w}_{\text{minvar}} = \arg\min_{\mathbf{w}} \mathbf{w}^\top \mathbf{\Sigma} \mathbf{w}$
    \item Maximum Sharpe portfolio: $\mathbf{w}_{\text{maxSharpe}} = \arg\max_{\mathbf{w}} \frac{\mu(\mathbf{w})}{\sigma(\mathbf{w})}$
  \end{itemize}

  \item \textbf{Mean vs Probability of Loss} ($\mu$--$p_{\text{loss}}$):
  Focuses on minimizing the chance of overvaluation. Identifies the portfolio with minimum $p_{\text{loss}}$ for a given expected return.

  \item \textbf{Mean vs Downside Deviation} ($\mu$--$\sigma_{\text{down}}$):
  Sortino-style frontier. Identifies the portfolio with maximum Sortino ratio:
  \[
  \text{Sortino}(\mathbf{w}) = \frac{\mu(\mathbf{w})}{\sigma_{\text{down}}(\mathbf{w})}
  \]

  \item \textbf{Mean vs CVaR} ($\mu$--$\text{CVaR}$):
  Tail risk frontier. Identifies the portfolio minimizing expected tail loss.

  \item \textbf{Mean vs VaR} ($\mu$--$\text{VaR}$):
  Similar to CVaR but focuses on the quantile rather than the conditional expectation.

  \item \textbf{Mean vs Maximum Drawdown} ($\mu$--$\text{MDD}$):
  Calmar-style frontier. Identifies the portfolio with maximum Calmar ratio:
  \[
  \text{Calmar}(\mathbf{w}) = \frac{\mu(\mathbf{w})}{\text{MDD}(\mathbf{w})}
  \]
\end{enumerate}

In each plot, the efficient frontier curve and key labeled portfolios (e.g., maximum return, minimum risk, optimal ratio) are annotated. Portfolio composition legends show the top 15 holdings for each optimal portfolio.

\subsubsection{Interpretation}

The probabilistic DCF framework yields a summary of valuation statistics and qualitative signals:

\begin{itemize}
  \item \textbf{Strong buy}: if both models (FCFE and FCFF) yield undervaluation ($\mathbb{P}[\text{Surplus} > 0] > 0.6$).
  \item \textbf{Buy}: if only one model suggests undervaluation.
  \item \textbf{Avoid}: if both suggest overvaluation ($\mathbb{P}[\text{Surplus} < 0] > 0.6$).
  \item \textbf{Hold}: if both suggest fair valuation.
\end{itemize}

The probabilistic approach provides several advantages over deterministic DCF:

\begin{enumerate}
  \item \textbf{Uncertainty quantification}: Rather than a single IVPS point estimate, we obtain a full distribution $\hat{f}(v)$, allowing probabilistic statements about valuation.

  \item \textbf{Confidence calibration}: The probability of undervaluation/overvaluation provides a confidence measure for investment decisions.

  \item \textbf{Portfolio-level analysis}: By combining simulation matrices across assets, we can construct efficient portfolios based on intrinsic value rather than historical prices.

  \item \textbf{Regime-invariant correlations}: The low inter-asset correlations (~0.0) observed in DCF-based returns reflect fundamental independence, unlike market-based returns which exhibit regime-dependent correlation structures.

  \item \textbf{Tail risk visibility}: The CVaR and maximum drawdown frontiers explicitly address worst-case scenarios, aligning with investor risk preferences during stress.
\end{enumerate}

The framework enables the construction of valuation-based portfolios that systematically exploit mispricing while managing fundamental uncertainty.

\newpage\subsection{GARP / PEG Analysis}

Growth at a Reasonable Price (GARP) bridges value and growth investing by seeking companies with solid growth prospects trading at reasonable valuations. The PEG ratio is the primary metric.

\subsubsection{PEG Ratio}

The Price/Earnings-to-Growth ratio normalizes valuation by growth rate:
\[
\text{PEG} = \frac{\text{P/E}}{\text{Growth Rate} \times 100}
\]

where Growth Rate is expressed as a decimal (e.g., 15\% = 0.15). A PEG of 1.0 implies fair value; below 1.0 suggests undervaluation relative to growth.

\paragraph{Growth Rate Selection}

The model selects growth rate with the following priority:
\begin{enumerate}
  \item 1-year EPS growth (most recent)
  \item Revenue growth (fallback)
  \item Earnings growth (secondary fallback)
  \item 5-year analyst estimate (last resort)
\end{enumerate}

All growth rates are capped at 50\% to prevent distortion from outliers.

\paragraph{PEGY Ratio}

For dividend-paying stocks, the dividend-adjusted PEG accounts for yield:
\[
\text{PEGY} = \frac{\text{P/E}}{\text{Growth Rate}(\%) + \text{Dividend Yield}(\%)}
\]

\paragraph{Fair Value Estimation}

Assuming a fair PEG of 1.0:
\[
\text{Fair P/E} = \text{Growth Rate} \times 100
\]
\[
\text{Fair Price} = \text{Fair P/E} \times \text{EPS}
\]

\subsubsection{GARP Scoring}

The composite GARP score (0--100) aggregates five components:

\begin{center}
\begin{tabular}{lcc}
\textbf{Component} & \textbf{Weight} & \textbf{Criteria} \\
\hline
PEG Score & 30 & PEG $< 0.5$: 30; $< 1.0$: 25; $< 1.5$: 15; $< 2.0$: 5 \\
Growth Score & 25 & $> 25\%$: 25; $> 15\%$: 20; $> 10\%$: 15; $> 5\%$: 10 \\
FCF Conversion & 20 & FCF/NI $> 1.0$: 20; $> 0.8$: 15; $> 0.5$: 10 \\
Balance Sheet & 15 & D/E $< 0.3$: 15; $< 0.5$: 10; $< 1.0$: 5 \\
ROE Score & 10 & $> 20\%$: 10; $> 15\%$: 7; $> 10\%$: 5 \\
\end{tabular}
\end{center}

\subsubsection{PEG Assessment}

\begin{itemize}
  \item[] PEG $< 0.5$: Very Undervalued
  \item[] PEG $0.5$--$1.0$: Undervalued
  \item[] PEG $1.0$--$1.5$: Fairly Valued
  \item[] PEG $1.5$--$2.0$: Moderately Expensive
  \item[] PEG $> 2.0$: Expensive
\end{itemize}

\newpage\subsection{Growth Analysis}

Growth analysis evaluates companies from a pure growth investing perspective, focusing on revenue acceleration, margin trajectory, and capital efficiency.

\subsubsection{Rule of 40}

The Rule of 40 is a benchmark for growth companies, particularly in technology and SaaS:
\[
\text{Rule of 40} = \text{Revenue Growth}(\%) + \text{Operating Margin}(\%)
\]

Companies exceeding 40 demonstrate efficient growth---either growing rapidly with acceptable margins or growing moderately with strong profitability.

\begin{itemize}
  \item[] $\geq 40$: Excellent
  \item[] $30$--$40$: Good
  \item[] $20$--$30$: Moderate
  \item[] $< 20$: Concerning
\end{itemize}

\subsubsection{Operating Leverage}

Operating leverage measures how earnings growth amplifies revenue growth:
\[
\text{Operating Leverage} = \frac{\text{Earnings Growth}}{\text{Revenue Growth}}
\]

Leverage $> 1$ indicates positive operating leverage (margins expanding with scale).

\subsubsection{EV/Revenue Efficiency}

Normalized valuation per unit of growth:
\[
\text{EV/Rev per Growth} = \frac{\text{EV/Revenue}}{\text{Revenue Growth}(\%)}
\]

Lower values indicate more attractive valuations relative to growth rate.

\subsubsection{Implied Growth Rate}

The market-implied growth rate from current valuation:
\[
g_{\text{implied}} = r - \frac{1}{\text{Forward P/E}}
\]

where $r$ is the required return (typically 10\%). This reveals what growth rate the market is pricing in.

\subsubsection{Growth Classification}

\begin{itemize}
  \item[] $> 40\%$: Hypergrowth
  \item[] $20$--$40\%$: High Growth
  \item[] $10$--$20\%$: Moderate Growth
  \item[] $5$--$10\%$: Slow Growth
  \item[] $< 5\%$: No Growth / Declining
\end{itemize}

\subsubsection{Growth Scoring}

The composite growth score (0--100) weights five factors:

\begin{center}
\begin{tabular}{lcc}
\textbf{Component} & \textbf{Weight} & \textbf{Key Thresholds} \\
\hline
Revenue Growth & 25 & $> 40\%$: 25; $> 20\%$: 18; $> 10\%$: 10 \\
Earnings Growth & 20 & $> 50\%$: 20; $> 30\%$: 16; $> 20\%$: 12 \\
Margin Quality & 20 & Gross $> 60\%$, Op $> 25\%$, FCF $> 20\%$ \\
Efficiency & 20 & Rule of 40, Operating Leverage \\
Quality (ROE/ROIC) & 15 & ROE $> 25\%$, ROIC $> 20\%$ \\
\end{tabular}
\end{center}

\newpage\subsection{Relative Valuation}

Relative valuation compares a target company against peer companies using standardized multiples, deriving fair value from comparable company analysis.

\subsubsection{Peer Similarity Scoring}

Peers are weighted by similarity across four dimensions (0--100 total):

\begin{center}
\begin{tabular}{lcc}
\textbf{Dimension} & \textbf{Weight} & \textbf{Scoring} \\
\hline
Industry & 30 & Same industry: 30; Same sector: 15 \\
Size & 25 & Market cap ratio $0.5$--$2.0\times$: 25 \\
Growth & 25 & Growth differential $< 5\%$: 25 \\
Profitability & 20 & EBITDA margin diff $< 5$pp: 20 \\
\end{tabular}
\end{center}

\subsubsection{Multiple Analysis}

For each valuation multiple $M$, we compute:

\paragraph{Premium/Discount}
\[
\text{Premium}(\%) = \frac{M_{\text{target}} - M_{\text{peer median}}}{M_{\text{peer median}}} \times 100
\]

\paragraph{Implied Valuation}

For equity multiples (P/E, P/B, P/S, P/FCF):
\[
\text{Implied Price} = \text{Metric} \times M_{\text{peer median}}
\]

For enterprise value multiples (EV/EBITDA, EV/Revenue):
\[
\text{Implied EV} = \text{Metric} \times M_{\text{peer median}}
\]
\[
\text{Implied Price} = \frac{\text{Implied EV} - \text{Debt} + \text{Cash}}{\text{Shares}}
\]

\subsubsection{Fair Premium Adjustment}

Companies with superior fundamentals deserve a premium:
\[
\text{Fair Premium} = \text{Growth Premium} + \text{Margin Premium}
\]

where:
\begin{itemize}
  \item[] Growth Premium: Each 5pp growth advantage $\rightarrow$ 10\% premium
  \item[] Margin Premium: Each 5pp margin advantage $\rightarrow$ 5\% premium
  \item[] Capped at: $-30\%$ to $+50\%$
\end{itemize}

\subsubsection{Relative Score}

The relative score (0--100) measures undervaluation:
\[
\text{Score} = 50 - \text{Average Premium}(\%)
\]

Higher scores indicate greater undervaluation relative to peers.

\subsubsection{Multiples Analyzed}

\begin{itemize}
  \item \textbf{Equity multiples}: P/E (trailing, forward), P/B, P/S, P/FCF
  \item \textbf{EV multiples}: EV/EBITDA, EV/EBIT, EV/Revenue
\end{itemize}

\newpage\subsection{Normalized Multiples}

Normalized multiples analysis compares a company's valuation multiples against sector benchmarks using percentile-based scoring with quality adjustments. Unlike relative valuation, which operates on a small peer group, this framework benchmarks against the full sector distribution.

\subsubsection{Multiple Extraction}

Ten valuation multiples are computed:

\paragraph{Price Multiples}

\begin{itemize}
  \item P/E (TTM, NTM): Price / Earnings per share
  \item P/S (TTM): Price / Revenue per share
  \item P/B (TTM): Price / Book value per share
  \item P/FCF (TTM): Price / Free cash flow per share
  \item PEG: P/E / Earnings growth rate
\end{itemize}

\paragraph{Enterprise Value Multiples}

\begin{itemize}
  \item EV/EBITDA (TTM): Enterprise value / EBITDA
  \item EV/EBIT (TTM): Enterprise value / EBIT
  \item EV/Sales (TTM): Enterprise value / Revenue
  \item EV/FCF (TTM): Enterprise value / Free cash flow
\end{itemize}

A multiple is considered valid when both the numerator and denominator are positive.

\subsubsection{Percentile Ranking}

Each multiple is ranked against the sector distribution using piecewise linear interpolation through three known quantiles ($Q_{25}$, $Q_{50}$, $Q_{75}$):

\[
\text{Pct}(x) = \begin{cases}
25 \cdot \dfrac{x}{Q_{25}} & \text{if } x \leq Q_{25} \\[8pt]
25 + 25 \cdot \dfrac{x - Q_{25}}{Q_{50} - Q_{25}} & \text{if } Q_{25} < x \leq Q_{50} \\[8pt]
50 + 25 \cdot \dfrac{x - Q_{50}}{Q_{75} - Q_{50}} & \text{if } Q_{50} < x \leq Q_{75} \\[8pt]
\min\!\left(100, \; 75 + 25 \cdot \dfrac{x - Q_{75}}{Q_{75} - Q_{50}}\right) & \text{if } x > Q_{75}
\end{cases}
\]

A percentile of 50 means the company trades at the sector median; lower percentiles indicate cheaper valuation.

\subsubsection{Premium/Discount and Implied Price}

\paragraph{Premium/Discount}

\[
\text{Premium}(\%) = \frac{M_{\text{company}} - M_{\text{sector median}}}{M_{\text{sector median}}} \times 100
\]

\paragraph{Implied Price}

For equity multiples (P/E, P/B, P/S, P/FCF):
\[
\text{Implied Price} = \text{Underlying Metric Per Share} \times M_{\text{sector median}}
\]

For enterprise value multiples (EV/EBITDA, EV/EBIT, EV/Sales, EV/FCF):
\[
\text{Implied Price} = \frac{(\text{Underlying Metric} \times M_{\text{sector median}}) - \text{Net Debt}}{\text{Shares Outstanding}} \times \frac{\text{Current Price}}{\text{Market Cap / Shares Outstanding}}
\]

\subsubsection{Quality Adjustment}

Companies with superior fundamentals deserve a premium; quality adjustment shifts the composite percentile accordingly:

\paragraph{Growth Premium}

\[
\text{Growth Premium}(\%) = \frac{g_{\text{company}} - g_{\text{sector median}}}{g_{\text{sector median}}} \times 10
\]

where $g$ is revenue growth (TTM). Clamped to $[-20\%, +20\%]$.

\paragraph{Margin Premium}

\[
\text{Margin Premium}(\%) = \frac{m_{\text{company}} - m_{\text{sector median}}}{m_{\text{sector median}}} \times 10
\]

where $m$ is the EBITDA margin. Clamped to $[-20\%, +20\%]$.

\paragraph{Return Premium}

\[
\text{Return Premium}(\%) = \frac{\text{ROE}_{\text{company}} - \text{ROE}_{\text{sector}}}{\text{ROE}_{\text{sector}}} \times 5
\]

Clamped to $[-20\%, +20\%]$.

\paragraph{Total Fair Premium}

\[
\text{Fair Premium}(\%) = \text{Growth} + \text{Margin} + \text{Return}
\]

\subsubsection{Composite Scoring}

\paragraph{Raw Composite Percentile}

The arithmetic mean of all valid multiples' percentile ranks:
\[
\overline{\text{Pct}} = \frac{1}{|\mathcal{M}|} \sum_{m \in \mathcal{M}} \text{Pct}(m)
\]

where $\mathcal{M}$ is the set of valid multiples.

\paragraph{Quality-Adjusted Percentile}

\[
\overline{\text{Pct}}_{\text{adj}} = \text{clamp}\!\left( \overline{\text{Pct}} - \text{Fair Premium}(\%), \; 0, \; 100 \right)
\]

A company deserving a 15\% quality premium has its percentile reduced by 15 points, making it appear cheaper on a quality-adjusted basis.

\paragraph{Confidence Score}

Confidence (0--100) combines data coverage and agreement:
\[
\text{Confidence} = \underbrace{\min(50, \; |\mathcal{M}| \times 5)}_{\text{Data score}} + \underbrace{\max(0, \; 50 - \sigma_{\text{Pct}})}_{\text{Agreement score}}
\]

where $\sigma_{\text{Pct}}$ is the standard deviation of percentile ranks across multiples. Low dispersion means multiples agree, yielding higher confidence.

\paragraph{Value Score}

The value score inverts the percentile:
\[
\text{Value Score} = 100 - \overline{\text{Pct}}_{\text{adj}}
\]

Higher value scores indicate deeper undervaluation after quality adjustment.

\subsubsection{Interpretation}

\begin{enumerate}
  \item \textbf{Sector context}: All comparisons are sector-relative; a ``cheap'' P/E in tech would be expensive in utilities. Sector benchmarks provide the appropriate reference frame.

  \item \textbf{Quality matters}: A company at the 80th percentile of P/E might be fairly valued if its growth and margins are also at the 80th percentile; the quality adjustment captures this.

  \item \textbf{Multiple agreement}: When P/E says cheap but EV/EBITDA says expensive, the confidence score drops, signaling caution.

  \item \textbf{Cyclicality}: Trailing multiples are distorted near cycle peaks and troughs. PEG and forward P/E provide partial correction.
\end{enumerate}

\newpage\subsection{Dividend Income Analysis}

Dividend income analysis evaluates stocks from an income investor's perspective, emphasizing yield sustainability, dividend growth, and payout safety.

\subsubsection{Dividend Discount Models}

\paragraph{Gordon Growth Model (Constant Growth DDM)}

For stocks with stable, perpetual dividend growth:
\[
\text{Fair Value} = \frac{D_1}{r - g} = \frac{D_0 (1 + g)}{r - g}
\]

where $D_0$ is the current dividend, $r$ is the required return, and $g$ is the dividend growth rate. Validity requires $r > g$.

\paragraph{Two-Stage DDM}

For companies transitioning from high growth to stable growth:
\[
\text{Fair Value} = \sum_{t=1}^{n} \frac{D_t}{(1+r)^t} + \frac{D_{n+1}}{(r - g_{\text{terminal}})} \cdot \frac{1}{(1+r)^n}
\]

\paragraph{H-Model (Gradual Growth Decline)}
\[
\text{Fair Value} = \frac{D_0 (1 + g_L)}{r - g_L} + \frac{D_0 \cdot H \cdot (g_S - g_L)}{r - g_L}
\]

where $H$ is the half-life of the high-growth period, $g_S$ is short-term growth, and $g_L$ is long-term sustainable growth.

\subsubsection{Dividend Coverage}

\[
\text{EPS Coverage} = \frac{\text{EPS}}{\text{DPS}}
\]
\[
\text{FCF Coverage} = \frac{\text{FCF per Share}}{\text{DPS}}
\]

Coverage $> 1.5\times$ is considered safe; $< 1.0\times$ indicates the dividend exceeds earnings.

\subsubsection{Chowder Number}

A metric for income-focused investors:
\[
\text{Chowder} = \text{Dividend Yield}(\%) + \text{Avg Dividend Growth}(\%)
\]

Values $> 12$--$14\%$ are attractive for total return from dividends.

\subsubsection{Dividend Aristocrat Classification}

\begin{itemize}
  \item[] \textbf{Dividend King}: 50+ years of consecutive increases
  \item[] \textbf{Dividend Aristocrat}: 25+ years
  \item[] \textbf{Dividend Achiever}: 10--24 years
  \item[] \textbf{Dividend Contender}: 5--9 years
  \item[] \textbf{Dividend Challenger}: 1--4 years
\end{itemize}

\subsubsection{Payout Ratio Assessment}

\begin{itemize}
  \item[] $< 50\%$: Very Safe
  \item[] $50$--$60\%$: Safe
  \item[] $60$--$75\%$: Moderate
  \item[] $75$--$90\%$: Elevated
  \item[] $> 90\%$: Unsustainable
  \item[] $> 100\%$: Paying from reserves
\end{itemize}

\subsubsection{Dividend Safety Scoring}

The dividend safety score (0--100) aggregates:

\begin{center}
\begin{tabular}{lcc}
\textbf{Component} & \textbf{Weight} & \textbf{Key Thresholds} \\
\hline
Payout Ratio & 25 & $< 50\%$: 25; $< 60\%$: 20; $< 75\%$: 15 \\
Coverage & 25 & EPS $> 1.5\times$: 25; $> 1.3\times$: 18 \\
Streak & 25 & Based on consecutive years \\
Balance Sheet & 15 & D/E $< 0.5$: 15; $< 1.0$: 10 \\
Stability & 10 & ROE/ROA quality metrics \\
\end{tabular}
\end{center}

\newpage\subsection{REIT Valuation}

Real Estate Investment Trusts require specialized valuation metrics due to significant non-cash depreciation charges and the asset-intensive nature of real estate.

\subsubsection{Funds From Operations (FFO)}

FFO is the REIT-standard earnings metric:
\[
\text{FFO} = \text{Net Income} + \text{Depreciation} + \text{Amortization} + \text{Impairments} - \text{Gains on Sales}
\]

FFO adds back non-cash real estate depreciation, which often exceeds economic depreciation for well-maintained properties.

\subsubsection{Adjusted FFO (AFFO)}

AFFO provides a more conservative view of recurring cash flow:
\[
\text{AFFO} = \text{FFO} - \text{Maintenance CapEx} - \text{Straight-Line Rent Adj} - \text{Stock Comp}
\]

AFFO better approximates true free cash flow available for distribution.

\subsubsection{Net Asset Value (NAV)}

NAV values the underlying real estate portfolio:
\[
\text{Property Value} = \frac{\text{NOI}}{\text{Cap Rate}}
\]
\[
\text{NAV} = \text{Property Value} + \text{Cash} + \text{Other Assets} - \text{Total Debt}
\]
\[
\text{NAV per Share} = \frac{\text{NAV}}{\text{Shares Outstanding}}
\]

\paragraph{Sector Cap Rates}

\begin{center}
\begin{tabular}{lclc}
\textbf{Sector} & \textbf{Cap Rate} & \textbf{Sector} & \textbf{Cap Rate} \\
\hline
Industrial & 5.0\% & Healthcare & 6.5\% \\
Data Center & 5.5\% & Office & 7.0\% \\
Self Storage & 5.5\% & Retail & 7.0\% \\
Residential & 5.0\% & Hotel & 8.0\% \\
\end{tabular}
\end{center}

\subsubsection{Valuation Multiples}

\[
\text{P/FFO} = \frac{\text{Price}}{\text{FFO per Share}}
\]
\[
\text{P/AFFO} = \frac{\text{Price}}{\text{AFFO per Share}}
\]
\[
\text{Premium/Discount to NAV} = \frac{\text{Price} - \text{NAV per Share}}{\text{NAV per Share}}
\]

\subsubsection{REIT Payout Analysis}

REITs must distribute $\geq 90\%$ of taxable income. The AFFO payout ratio indicates sustainability:
\[
\text{AFFO Payout Ratio} = \frac{\text{Annual Dividend}}{\text{AFFO per Share}}
\]

\begin{itemize}
  \item[] $< 75\%$: Very Safe (room for growth)
  \item[] $75$--$85\%$: Healthy
  \item[] $85$--$95\%$: Tight
  \item[] $> 95\%$: At Risk
\end{itemize}

\subsubsection{Mortgage REIT Metrics}

Mortgage REITs (mREITs) invest in mortgages rather than properties:
\[
\text{Net Interest Income} = \text{Interest Income} - \text{Interest Expense}
\]
\[
\text{Net Interest Margin} = \frac{\text{NII}}{\text{Average Earning Assets}}
\]
\[
\text{Price/Book} = \frac{\text{Price}}{\text{Book Value per Share}}
\]

For mREITs, P/B near 1.0 is typical; significant discounts may indicate credit concerns.

\newpage\subsection{ETF Analysis}

ETF analysis evaluates exchange-traded funds on cost efficiency, tracking quality, liquidity, and structural characteristics.

\subsubsection{Premium/Discount to NAV}

\[
\text{Premium/Discount} = \frac{\text{Price} - \text{NAV}}{\text{NAV}} \times 100
\]

ETFs should trade near NAV; persistent premiums or discounts indicate inefficiencies or structural issues.

\subsubsection{Tracking Error}

Tracking error measures deviation from the benchmark:
\[
\text{Tracking Error} = \sqrt{\text{Var}(R_{\text{ETF}} - R_{\text{Benchmark}})}
\]

\begin{itemize}
  \item[] $< 0.10\%$: Excellent
  \item[] $0.10$--$0.25\%$: Good
  \item[] $0.25$--$0.50\%$: Acceptable
  \item[] $> 0.50\%$: Poor
\end{itemize}

\subsubsection{Total Cost of Ownership}

Beyond expense ratio, total cost includes:
\[
\text{TCO} = \text{Expense Ratio} + \text{Bid-Ask Spread} + \text{Tracking Drag}
\]

\subsubsection{ETF Quality Scoring}

The ETF quality score (0--100) aggregates:

\begin{center}
\begin{tabular}{lcc}
\textbf{Component} & \textbf{Weight} & \textbf{Key Thresholds} \\
\hline
Cost & 25 & ER $< 0.05\%$: 25; $< 0.10\%$: 20; $< 0.20\%$: 15 \\
NAV Gap & 10 & Gap $< 0.05\%$: 10; $< 0.10\%$: 8; $< 0.25\%$: 6 \\
Tracking & 25 & TE $< 0.10\%$: 25; $< 0.25\%$: 20; $< 0.50\%$: 15 \\
Liquidity & 20 & Spread, volume, AUM \\
Size (AUM) & 20 & $> \$100$B: 20; $> \$10$B: 15; $> \$1$B: 10 \\
\end{tabular}
\end{center}

\subsubsection{Derivatives ETF Analysis}

Specialized metrics for options-based and leveraged ETFs:

\paragraph{Capture Ratios}

Capture ratios measure how much of the benchmark's return is captured in up and down markets:
\[
\text{Upside Capture} = \frac{\sum_{t \in U} R_{\text{ETF},t}}{\sum_{t \in U} R_{\text{bench},t}} \times 100, \qquad
\text{Downside Capture} = \frac{\sum_{t \in D} R_{\text{ETF},t}}{\sum_{t \in D} R_{\text{bench},t}} \times 100
\]

where $U = \{t : R_{\text{bench},t} > 0\}$ and $D = \{t : R_{\text{bench},t} < 0\}$.

Capture efficiency isolates the asymmetry:
\[
\text{Capture Efficiency} = \text{Upside Capture} - \text{Downside Capture}
\]

Positive efficiency means the ETF captures more upside than downside---desirable for covered call strategies.

\paragraph{Covered Call ETFs}

Covered call ETFs (e.g., JEPI, QYLD) sacrifice upside for income. Key metrics:
\begin{itemize}
  \item Distribution Yield: annualized income as \% of NAV
  \item Yield vs Benchmark: $\text{Yield}_{\text{ETF}} / \text{Yield}_{\text{benchmark}}$
  \item Scoring sweet spot: 7--10\% yield with upside capture $\geq 60\%$
\end{itemize}

\paragraph{Volatility ETF Decay}

Volatility ETFs that hold VIX futures suffer from contango roll yield:
\[
\text{Roll Yield (monthly)} \approx \frac{R_{\text{1Y}}}{12}
\]

where $R_{\text{1Y}}$ is the 1-year return. Persistent contango (front futures $<$ back futures) causes compounding erosion. An ETF with $R_{\text{1Y}} < -30\%$ is flagged as severely decaying and unsuitable for buy-and-hold.

\paragraph{Leveraged ETF Volatility Drag}

Daily-rebalanced leveraged ETFs suffer from volatility drag:
\[
R_{\text{leveraged}} \approx L \cdot R_{\text{index}} - \frac{L(L-1)}{2} \sigma^2
\]

where $L$ is the leverage factor and $\sigma$ is realized volatility. Higher volatility accelerates decay; ETFs with $\sigma_{\text{1Y}} > 50\%$ score lowest.

\paragraph{Buffer ETFs}

Buffer (defined-outcome) ETFs use options collars to provide downside protection up to a buffer level in exchange for a capped upside:
\begin{itemize}
  \item Buffer Level: starting point of downside protection
  \item Cap Level: maximum upside participation
  \item Remaining Buffer / Cap: current effective levels within the outcome period
  \item Days to Outcome Period Reset
\end{itemize}

Mid-period entry alters effective buffer and cap levels; best entered near period start.

\newpage\subsection{Crypto Treasury Valuation}

Crypto treasury valuation applies to companies holding significant cryptocurrency assets on their balance sheet, using a multiple of Net Asset Value (mNAV) framework.

\subsubsection{Net Asset Value}

\[
\text{NAV} = (\text{BTC Holdings} \times P_{\text{BTC}}) + (\text{ETH Holdings} \times P_{\text{ETH}})
\]
\[
\text{NAV per Share} = \frac{\text{NAV}}{\text{Shares Outstanding}}
\]

\subsubsection{Multiple of NAV (mNAV)}

The key valuation metric:
\[
\text{mNAV} = \frac{\text{Market Cap}}{\text{NAV}}
\]

\begin{itemize}
  \item[] mNAV $< 0.8$: Strong Buy (20\%+ discount to holdings)
  \item[] mNAV $0.8$--$1.0$: Buy (discount)
  \item[] mNAV $1.0$--$1.2$: Hold (fair to slight premium)
  \item[] mNAV $1.2$--$1.5$: Caution (meaningful premium)
  \item[] mNAV $> 1.5$: Overvalued (speculative premium)
\end{itemize}

\subsubsection{Premium/Discount}

\[
\text{Premium}(\%) = (\text{mNAV} - 1) \times 100
\]

Negative values indicate the stock trades at a discount to its crypto holdings.

\subsubsection{Per-Share Exposure}

\[
\text{BTC per Share} = \frac{\text{BTC Holdings}}{\text{Shares Outstanding}}
\]

This metric shows the crypto exposure obtained per share purchased.

\subsubsection{Implied Crypto Prices}

The market-implied crypto price embedded in the stock valuation:
\[
\text{Implied BTC Price} = P_{\text{BTC}} \times \text{mNAV}
\]

If mNAV $> 1$, the market implicitly values the company's BTC holdings above spot price.

\subsubsection{Unrealized Gains}

\[
\text{Unrealized Gain} = (P_{\text{current}} - P_{\text{avg cost}}) \times \text{Holdings}
\]

Significant unrealized gains may create tax overhang if liquidated.

\newpage\subsection{NLP Sentiment Analysis}

Natural language processing analyzes corporate communications to detect narrative shifts before they manifest in financial results.

\subsubsection{Document Sources}

\begin{itemize}
  \item \textbf{SEC Filings}: MD\&A (Item 7/Item 5), Risk Factors (Item 1A/Item 3D)
  \item \textbf{Earnings Calls}: Prepared remarks, Q\&A transcripts
  \item \textbf{Comparison}: Current vs prior quarter, current vs historical centroid
\end{itemize}

\subsubsection{Embedding-Based Change Detection}

Documents are converted to dense vector embeddings using transformer models. Change is measured via cosine distance:
\[
\Delta = 1 - \cos(\mathbf{e}_{\text{current}}, \mathbf{e}_{\text{prior}})
\]

Higher $\Delta$ indicates greater semantic shift in language.

\subsubsection{Historical Centroid}

A weighted average of past embeddings with exponential decay:
\[
\mathbf{c} = \frac{\sum_{t} w_t \cdot \mathbf{e}_t}{\sum_{t} w_t}, \quad w_t = e^{-\lambda (T - t)}
\]

where $\lambda$ controls decay speed and $T$ is the current period. Recent quarters receive higher weight.

\subsubsection{Hedging Language Detection}

Count of uncertainty-indicating words and phrases:
\begin{itemize}
  \item \textbf{Hedging words}: ``may'', ``might'', ``could'', ``uncertain'', ``risk''
  \item \textbf{Evasive phrases}: ``as I mentioned'', ``too early to say'', ``time will tell''
\end{itemize}

Increasing hedging language signals growing management uncertainty.

\subsubsection{Commitment Strength Scoring}

Management commitment is scored on a ladder (0--1):

\begin{center}
\begin{tabular}{lclc}
\textbf{Word} & \textbf{Score} & \textbf{Word} & \textbf{Score} \\
\hline
``will'' & 1.0 & ``intend'' & 0.5 \\
``expect'' & 0.8 & ``target'' & 0.4 \\
``anticipate'' & 0.7 & ``hope'' & 0.3 \\
``believe'' & 0.6 & ``monitor'' & 0.2 \\
\end{tabular}
\end{center}

Declining commitment scores indicate weakening conviction.

\subsubsection{Signal Ontology}

Fixed categories for narrative signals (each with UP/DOWN direction):

\begin{center}
\begin{tabular}{ll}
\textbf{Category} & \textbf{Interpretation} \\
\hline
Pricing Power & Ability to raise/maintain prices \\
Margin Pressure & Cost/competitive pressure on margins \\
Demand Visibility & Clarity on demand outlook \\
Competitive Threat & New or intensifying competition \\
Guidance Confidence & Management confidence in outlook \\
Management Tone & Overall tone shift \\
Regulatory Overhang & Regulatory risk exposure \\
Execution Risk & Ability to execute initiatives \\
\end{tabular}
\end{center}

\subsubsection{Composite Scoring}

Snippets are ranked by composite score:
\[
\text{Score} = w_{\Delta} \cdot \Delta_{\text{embedding}} + w_h \cdot \Delta_{\text{hedging}} + w_c \cdot \Delta_{\text{commitment}}
\]

where typical weights are $w_{\Delta} = 0.5$, $w_h = 0.25$, $w_c = 0.25$.

\subsubsection{Interpretation}

\begin{itemize}
  \item \textbf{High $\Delta$ + Decreasing Commitment}: Deteriorating narrative
  \item \textbf{Increasing Hedging}: Growing uncertainty/caution
  \item \textbf{Risk Factor Changes}: New risks emerging
  \item \textbf{Q\&A Tone Shift}: Management becoming defensive
\end{itemize}

\newpage\subsection{Tail Risk Forecast}

Tail risk forecasting estimates Value-at-Risk (VaR) and Expected Shortfall (ES) using realized variance computed from high-frequency intraday returns.

\subsubsection{Realized Variance}

Daily realized variance is computed from intraday returns:
\[
RV_t = \sum_{i=1}^{n} r_{t,i}^2
\]

where $r_{t,i}$ are intraday log returns (typically 5-minute intervals). This provides a more accurate volatility measure than close-to-close returns by capturing intraday price movements.

\subsubsection{HAR-RV Model}

The Heterogeneous Autoregressive Realized Volatility (HAR-RV) model captures volatility persistence at multiple horizons:
\[
RV_{t+1} = c + \beta_d \cdot RV_t + \beta_w \cdot RV_t^{(w)} + \beta_m \cdot RV_t^{(m)} + \varepsilon_t
\]

where:
\begin{itemize}
  \item $RV_t^{(w)} = \frac{1}{5}\sum_{i=0}^{4} RV_{t-i}$ is the weekly (5-day) average
  \item $RV_t^{(m)} = \frac{1}{22}\sum_{i=0}^{21} RV_{t-i}$ is the monthly (22-day) average
\end{itemize}

The HAR structure captures heterogeneous market participants operating at different frequencies (daily traders, weekly rebalancers, monthly allocators).

\subsubsection{Jump Detection}

Variance jumps are identified using a threshold approach:
\[
J_t = \mathbf{1}\left\{ RV_t > \mu_{RV} + k \cdot \sigma_{RV} \right\}
\]

where $\mu_{RV}$ and $\sigma_{RV}$ are the rolling mean and standard deviation of realized variance, and $k = 2.5$ is the threshold multiplier. When recent jumps are detected, the forecast includes a risk premium adjustment.

\subsubsection{Value-at-Risk}

VaR at confidence level $\alpha$ estimates the maximum loss that will not be exceeded with probability $\alpha$:
\[
\text{VaR}_\alpha = t_\alpha(\nu) \cdot \sigma_{\text{forecast}}
\]

where $t_\alpha(\nu)$ is the quantile of the Student-t distribution with $\nu = 6$ degrees of freedom, accounting for fat tails in return distributions.

\begin{center}
\begin{tabular}{lcc}
\textbf{Confidence} & \textbf{t(6) Quantile} & \textbf{Interpretation} \\
\hline
95\% & 1.943 & 5\% chance of exceeding \\
99\% & 3.143 & 1\% chance of exceeding \\
\end{tabular}
\end{center}

\subsubsection{Expected Shortfall}

Expected Shortfall (Conditional VaR) estimates the average loss in the worst scenarios:
\[
\text{ES}_\alpha = \text{VaR}_\alpha \times m_\alpha
\]

where the multiplier $m_\alpha$ accounts for the tail weight:
\begin{itemize}
  \item $m_{0.95} = 1.25$ for 95\% confidence
  \item $m_{0.99} = 1.35$ for 99\% confidence
\end{itemize}

ES is more conservative than VaR and better captures tail risk for non-normal distributions.

\subsubsection{Interpretation}

\begin{center}
\begin{tabular}{ll}
\textbf{Metric} & \textbf{Meaning} \\
\hline
VaR 95\% = 2\% & 5\% chance of losing more than 2\% tomorrow \\
VaR 99\% = 3.5\% & 1\% chance of losing more than 3.5\% tomorrow \\
ES 95\% = 2.5\% & Average loss in worst 5\% of days is 2.5\% \\
Jump-adjusted & Recent variance jumps detected; forecast conservative \\
\end{tabular}
\end{center}

Higher volatility stocks (e.g., NVDA: VaR 99\% $\approx$ 7.5\%) require larger position sizing buffers than lower volatility names (e.g., SPY: VaR 99\% $\approx$ 2.3\%).
